Diversification of Australian blind snakes

### 2022-12-09 manuscript.Rmd
update discussion

Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10). The model indicates two pulses of biogeographic transitions for the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a. The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, but 12/21 lineages expanded or transitioned into other biomes. The common ancestor of *Group 2* transitioned into the arid biome, and 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.

@@

Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by multiple transitions into the arid biome and subsequent expansion to the Mediterranean biome. However, our results show that the timing of these transition and diversification in the arid biome, for both species in groups 1 and 2, occured earlier (20 – 10 MYA) than previously proposed (10 MYA) by @marinTracing2013.

@@

Alternatively, this suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy red centre of Australia [@byrneBirth2008]. 
This pattern may indicate that blind snakes, like other fossorial animals, are pre-adapted to arid environments (CITE). Subterranean microhabitat is relatively stable compared to ground.

>>>>

Based on current distributions (ALA) most _Anilios_ species are found entirely or at least partially in the tropical grassland biome (n = 20) followed by the arid biome (n = 13) and mediterranean biome (n = 10). The model indicates two pulses of biogeographic transitions for the two major clades, which we will refer to as *Group 1* (n=21 lineages) and *Group 2* (n=14 lineages) (\@ref(fig:fig-biogeobears)a. The common ancestor of *Group 1* remained mostly in the Tropical Grassland Biome, then 12/21 lineages expanded or transitioned into other biomes. The common ancestor of *Group 2* first transitioned into the arid biome, and then 7/14 lineages expanded to the Mediterranean and Temperate Forest biomes, and/or transitioned back into the Tropical Grassland Biome.

@@

Consistent with @marinTracing2013, we found the initial diversification occurred very rapidly (<1 million years) in the tropical grassland biome followed by transitions and/or expansion into other biomes. 
However, as opposedrecent invasion (<10 MYA) and diversification (<5 MYA) into the arid zone [@marinHidden2013], our results show that species in both groups 1 and 2 have deep divergence between 20 – 10 MYA, during the time when the Australia was still relatively wet and humid [@byrneBirth2008]. This suggests that these lineages have diversified before the intensive continent-wide aridification in the late Miocene that resulted in the present dry and sandy areas of the arid biome [@byrneBirth2008].

One hypothesis to explain contemporary distribution in the arid biome is that *Anilios* may be pre-adapted to arid environments. Blind snakes have been found in xeric areas around the world, for example _Myriopholis macrorhyncha_ in the desert

North American  and the Palearctic _Letheobia episcopus_) possibly because subterranean microhabitat has relatively stable climate compared to surface [@kinlawReview1999;@nevoAdaptive1979].

Australian blind snakes and these blind snakes share similar morphological features and diet consisting of eggs, larvae, and pupae. of ants and termites.

The fact that we observed multiple independent transitions into the arid biome.

@@
